Open Data

### 2020-05-29 draft_chapter.Rmd
Final edits

```{r setup, include=FALSE, eval = F}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
# library(jsonlite)
# library(osmdata)
# library(ggplot2)
# library(sf)
# library(readr)
# library(tidyr)
# library(patchwork)
# library(dplyr)
```

@@

The 'public sector' refers to organizations owned and operated by local or federal government with the core aim to provide services to the public. Much of the open data movement has focused on opening up information generated by these local and national state agencies [@kitchin2014data]. These can include national statistics or administrative data, but also data collected by publicly funded research projects such as victimization and household surveys, amongst others. The _Global Open Data Index_, introduced earlier, focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].

@@

So what can we do to make sure we make good use of these data? With a critical, engaged and considered approach to conducting research with open data, criminology can become a leading force in open and reproducible social science. In sum, researchers and analysts using open data must first ask critical questions:

@@

In the United Kingdom, police recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence. Visit this website, and you will see a welcome message, and six tabs across the top of it, which should read "Home", "Data", "API", "Changelog", "Contact", and "About".

Before we download any data, we can learn more about it by clicking on the "About" tab. This brings up a lot of information that it is important to review carefully in order to answer the where, why, who, what and when questions posed earlier. Take a moment to read through this information, and try to take notes on what you think might be relevant for your analysis. For example, if we want to map crimes, we might want to explore if there is any type of "anonymization" that might take place before the data are released (to protect the privacy of the victims). If you read the "About" page, you might find the following note:

> Location anonymization
The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened.

This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate/ This may have implications for our findings later!

To then actually download some data, move on to clicking on the "Data" tab. This should open a page entitled "Data downloads" under which you can see another five tabs: "Custom download", "Archive", "Boundaries", "Open data", and "Statistical data".  By staying on this "Custom download" page you can select what sort of data you want to download. We can select the time period and police force of interest, and the type of information required (e.g. crimes, stop and search, outcomes). We are also informed that the data are downloaded in comma-separated values format (.csv file extension), which meets our machine-readable, easy-to-manipulate data format requirements.

@@

Save this file locally (in your working directory) to a subfolder named "data". You can load it in with the `read_csv()` command from the `readr` package. Remember that you will need to load it with `library(readr)`.

```{r crimeload1}
library(readr)
btp_df <- read_csv("data/2020-01-btp-street.csv")
```

Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our github page. Note that the line break below is purely for easy formatting - this will need removing in your own script so that the URL runs across one line.

@@

Here, we will make a direct request to the API created by London's local government organization responsible for transport: Transport for London (TfL). TfL oversee the London Underground. They provide access to their open data through a unified API. Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What', 'When' questions, we must find a similar document for TfL. Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.

They further provide a documentation page with examples of how you can use HTTP (i.e. web link) requests to make calls to the API (see: https://api-portal.tfl.gov.uk/docs). If you follow this link you should see a page titled "Our Unified API" and some examples of calls.

To draw conclusions about crimes in and around London Underground stations, we will need some spatial data about the stations themselves.

In the documentation, there is a section called 'API area' which offer some guidance on the types of data available. Scroll down and find the example called _Stops_. You will see that this call returns information on stops for buses or London Underground lines, and there are two examples (bus route 24 and the Bakerloo line). Copy their demonstration URL for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser. When you visit this page you should see something like Figure 1.

@@

This gives us an object (`api_call`) which contains all the information returned by the TfL API. JSON is slightly different to traditional data frames with rows and columns, which we are probably more familiar with, because the data is nested. For instance, `api_call` is classed as a data frame, but now try viewing the object using `View(api_call)`. You will notice that some of the columns are actually lists, rather than character or factor vectors, which what we might usually expect. Another way of exploring the `api_call` object is by looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.

This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated. It is not always a neatly formatted data frame like the csv from the open police data portal. That said, we can transform this data into something more familiar from within R, as we will see later on.

@@

If we load the package `osmdata` we can use its functions to query the Open Street Map API, rather than the API query being made directly (like we did for TfL, using the URL). Once again, we will want to identify documentation which can help us understand and critique our data, and learn how to query it. When it comes to OSM, you can read all about it on their community page: [https://www.openstreetmap.org/about](https://www.openstreetmap.org/about). Using the wrapper in R, we can refer to the package documentation and the associated vignette which we can find at [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).

Unlike the TfL API, `osmdata` is an international database, and has lots of data that we might not necessarily need. As such, the first thing we need to specify is our study region. To do this, we use the `getbb()` function from the `osmdata` package, which stands for "get bounding box". You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the OSM API that we want everything *inside* the box, but nothing *outside* the box.

So, how can we name a bounding box specification to define the study region? This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, or you can use a search term. Here, we want to select Greater London, so we can use the search term "greater london united kingdom". Besides specifying the study region, we can also tell the `getbb()` function what format we want the data to be in. In this case we want a spatial object, specifically an `sf` polygon object, which we name `bb_sf`:

@@

We now have an `sf` object with all the public transport stations in London mapped by OSM volunteers, along with 130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others. Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not have added information. Nevertheless, when the details are recorded, they provide rich insight and local knowledge we may otherwise not be able to obtain.

@@

We usually want to work with data that is in the format of a data frame, that is, all your observations are rows and all your variables are columns. While this is the case for our police data and our OSM data, recall that the TfL data came in the nested JSON format.

@@

We can now plot the OSM station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two data sets compare. We do not plot the study region boundary in order to get a more detailed, local view.

@@

As we can see, there is clearly some discrepancy between the two data sources. There are slight differences in the exact point locations of stations, but more saliently,  Open Street Map appears to be missing stations, especially in Central and South London. This could be due to a number of reasons, among them: variation in residents' willingness to contribute to information about their local area, and the demographic characteristics known to influence the volume of contributions [@mashhadi2013putting]. Given that South London is historically less wealthy in comparison to the North, for instance, this would certainly be an interesting avenue to pursue. Interestingly, these missing stations _have_ actually been recorded by contributors in Open Street Map, but _without_ information about the line on which the station is located. We can see this by plotting out all Open Street Map stations using `osm_stat_sf`. That said, without local knowledge, an alternative open data source, or a critical approach to using open data, well-intentioned researchers might draw erroneous conclusions from filtering stations tagged as on the Northern line. This gives us a preliminary answer to our earlier questions: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.

But, what might the impact of this be on studying crime in and around London Underground stations, given the differences observed between the two open data sources? For that, we can turn to our British Transport Police data.

@@

We now have the location of all crimes recorded by the British Transport Police in Greater London during January, 2020. For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each. It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police recorded crime data [see tompson2015uk], but it does serve to facilitate this demonstration.

@@

```{r crimemapcompare}
library(patchwork)

p1 <- ggplot(data = osm_north_sf) +
  geom_sf(mapping = aes(color = crimes), size = 2) +
  labs(title = "Open Street Map") +
  scale_color_viridis_c()

@@

@@

As showcased, the resulting visualizations are highly similar (a Spatial Point Pattern Test will confirm this, see Chapter 8, Topic 4). Refer back to the example earlier if some of the code is unclear. Note that this is a major benefit of using an API wrapper from within R, because the code used to query each open data source can easily be re-used, with minor amendments, to extract different features.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
# library(jsonlite)
# library(osmdata)
# library(ggplot2)
# library(sf)
# library(readr)
# library(tidyr)
# library(patchwork)
# library(dplyr)
```

@@

The 'public sector' refers to organizations owned and operated by local or central government with the core aim to provide services to the public. Much of the open data movement has focused on opening up information generated by these local and national state agencies [@kitchin2014data]. These can include national statistics or administrative data, but also data collected by publicly funded research projects such as victimization and household surveys, amongst others. The _Global Open Data Index_, introduced earlier, focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].

@@

So, what can we do to make sure we make good use of these data? With a critical, engaged and considered approach to conducting research with open data, criminology can become a leading force in open and reproducible social science. In sum, researchers and analysts using open data must first ask critical questions:

@@

In the United Kingdom, police recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence. Visiting this website, you will see a welcome message, and six tabs across the top which should read "Home", "Data", "API", "Changelog", "Contact", and "About".

Before we download any data, we can learn more about it by clicking on the "About" tab. This brings up information that is important to review carefully in order to answer the where, why, who, what and when questions posed earlier. Take a moment to read through this information, and take notes on what you think might be relevant for your analysis. For example, if we want to map crimes, we will want to explore if there is any type of "anonymization" that might take place before the data are released (to protect the privacy of the victims). If you read the "About" page, you might find the following note:

> Location anonymization.
The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened.

This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate. This may have implications for our findings!

To then download some data, move on to clicking on the "Data" tab. This should open a page entitled "Data downloads" under which you can see another five tabs: "Custom download", "Archive", "Boundaries", "Open data", and "Statistical data".  By staying on this "Custom download" page you can select what sort of data you want to download. We can select the time period and police force of interest, and the type of information required (e.g. crimes, stop and search, outcomes). We are also informed that the data are downloaded in comma-separated values format (.csv file extension), which meets our machine-readable, easy-to-manipulate data format requirements.

@@

Save this file locally (in your working directory) to a subfolder named "data". You can load it in with the `read_csv()` command from the `readr` package. Remember that you will need to load it first using `library()`.

```{r crimeload1}
library(readr)

btp_df <- read_csv("data/2020-01-btp-street.csv")
```

Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our GitHub page. Note that the line break below is purely for easy formatting - this will need removing in your own script so that the URL runs across one line.

@@

Here, we will make a direct request to the API created by London's local government organization responsible for transport: Transport for London (TfL). TfL oversee the London Underground. They provide access to their open data through a unified API. Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What' and 'When' questions, we must find a similar document for TfL. Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.

They further provide a documentation page with examples of how you can use HTTP (i.e. web link) requests to make calls to the API (see: https://api-portal.tfl.gov.uk/docs). If you follow this link you should see a page titled "Our Unified API" and some examples of calls. To draw conclusions about crimes in and around London Underground stations, we will need some spatial data about the stations themselves.

In the documentation, there is a section called 'API area' which offers some guidance on the types of data available. Scroll down and find the example called _Stops_. You will see that this call returns information on stops for buses or London Underground lines, and there are two examples (bus route 24 and the Bakerloo line). Copy their demonstration URL for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser. When you visit this page, you should see something like Figure 1.

@@

This gives us an object (`api_call`) which contains all the information returned by the TfL API. JSON is slightly different to traditional data frames with rows and columns, which we are probably more familiar with, because the data is nested. For instance, `api_call` is classed as a data frame, but now try viewing the object using `View(api_call)`. You will notice that some of the columns are actually lists, rather than character or factor vectors, which is what we might usually expect. Another way of exploring the `api_call` object is by looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.

This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated. It is not always a neatly formatted data frame like the CSV file from the open police data portal. That said, we can transform this data into something more familiar from within R, as we will see later on.

@@

If we load the package `osmdata` we can use its functions to query the Open Street Map API, rather than the API query being made directly (like we did for TfL, using the URL). Once again, we will want to identify documentation which can help us understand and critique our data, and learn how to query it. When it comes to Open Street Map, you can read all about it on their community page: [https://www.openstreetmap.org/about](https://www.openstreetmap.org/about). Using the wrapper in R, we can refer to the package documentation and the associated vignette online: [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).

Unlike the TfL API, `osmdata` is an international database, and has lots of data that we might not necessarily need. As such, the first thing we need to specify is our study region. To do this, we use the `getbb()` function from the `osmdata` package, which stands for "get bounding box". You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the Open Street Map API that we want everything *inside* the box, but nothing *outside* the box.

So, how can we name a bounding box specification to define the study region? This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, or you can use a search term. Here, we want to select Greater London, so we can use the search term "greater london united kingdom". Besides specifying the study region, we can also tell the `getbb()` function what format we want the data to be in. In this case we want a spatial object, specifically an `sf` polygon object, which we name `bb_sf`.

@@

We now have an `sf` object with all the public transport stations in London mapped by Open Street Map volunteers, along with 130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others. Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not added information. Nevertheless, when the details are recorded, they provide rich insight and local knowledge that we may otherwise be unable to obtain.

@@

We usually want to work with data that is in the format of a data frame, that is, all your observations are rows and all your variables are columns. While this is the case for our police data and our Open Street Map data, recall that the TfL data came in the nested JSON format.

@@

We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two data sets compare. We do not plot the study region boundary in order to get a more detailed, local view.

@@

As we can see, there is clearly some discrepancy between the two data sources. There are slight differences in the exact point locations of stations, but more saliently,  Open Street Map appears to be missing stations, especially in Central and South London. This could be due to a number of reasons, among them: variation in residents' willingness to contribute to information about their local area, and the demographic characteristics known to influence the volume of contributions [@mashhadi2013putting]. Interestingly, these missing stations _have_ actually been recorded by contributors in Open Street Map, but _without_ information about the line on which the station is located. We can see this by plotting out all Open Street Map stations using `osm_stat_sf`. That said, without local knowledge, an alternative open data source, or a critical approach to using open data, well-intentioned researchers might draw erroneous conclusions from filtering stations tagged as on the Northern line. This gives us a preliminary answer to our earlier question: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.

But, what might the impact of this be when studying crime in and around London Underground stations, given the differences observed between the two open data sources? For that, we can turn to our British Transport Police data.

@@

We now have the recorded location of all crimes recorded by the British Transport Police in Greater London during January, 2020. For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each. It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police recorded crime data [see @tompson2015uk], but it does serve to facilitate this demonstration.

@@

```{r crimemapcompare}
p1 <- ggplot(data = osm_north_sf) +
  geom_sf(mapping = aes(color = crimes), size = 2) +
  labs(title = "Open Street Map") +
  scale_color_viridis_c()

@@

## Alternative example

@@

As showcased, the resulting visualizations are highly similar (a Spatial Point Pattern Test will confirm this, see Chapter 8, Topic 4). Refer back to the example earlier if some of the code is unclear. Note that this is a major benefit of using an API wrapper from within R, because the code used to query each feature can easily be re-used, with minor amendments, to extract different information.
